---
output:
  html_document:
    keep_md: true
  pdf_document: default
---

# Report
## Subtitle: Practicum II CS5200
#### Author: Dhivya Kumaraguruparan
#### Date: Spring 2024


```{r label= "dbConnection", eval = T, warning = F, echo=FALSE, message= FALSE}
# 1. Library
library(RMySQL)
library(DBI)
library(RSQLite)
library(knitr)

# 2. Settings
db_user <- 'dhivya'            #  value from the setup
db_password <- 'Practicum@2024'    #  value from the setup
db_name <- 'practicum12024'         #  value from the setup
db_host <- 'db4free.net'       # for db4free.net
db_port <- 3306

# 3. Connect to remote server database
con <-  dbConnect(RMySQL::MySQL(), user = db_user, password = db_password,
                   dbname = db_name, host = db_host, port = db_port)

```


## Analytical Query I
### Top five sales reps with the most sales broken down by year

Based on their overall sales performance, the research determines the top five sales reps for each year. 2020 saw Walison da Silva lead the group, closely followed by Aneeta Kappoorthy and Lynette McRowe. In 2021 and 2022, the pattern persisted, with reliable performers topping the charts. The business can use the success of these top performers as leverage to increase overall sales growth by identifying and rewarding them. Frequent oversight and assistance can boost their output even further and offer insightful information that can be used to raise the performance of the sales force as a whole.

```{sql connection=con, label= "Top five sales reps with the most sales broken down by year sql", eval = T, warning = F, echo=FALSE, message= FALSE}
SELECT
    rep_name,
    SUM(total_amount_sold) AS total_sales,
    sale_year
FROM
    rep_facts
GROUP BY
    rep_name, sale_year
ORDER BY
    sale_year, total_sales DESC


```


```{r label= "Top five sales reps with the most sales broken down by year", eval = T, warning = F, echo=FALSE, message= FALSE}

# Check if ggplot2 package is installed, if not, install it
if (!require("ggplot2", quietly = TRUE)) {
  install.packages("ggplot2")
}

# Check if plotly package is installed, if not, install it
if (!require("plotly", quietly = TRUE)) {
  install.packages("plotly")
}

library(ggplot2)
library(plotly)

# SQL query
query <- "
    SELECT
        rep_name,
        SUM(total_amount_sold) AS total_sales,
        sale_year
    FROM
        rep_facts
    GROUP BY
        rep_name, sale_year
    ORDER BY
        sale_year, total_sales;
"

# Execute the query and store the result in 'result'
result <- DBI::dbGetQuery(con, query)

# Convert 'sale_year' to factor to maintain the order in the plot
result$sale_year <- factor(result$sale_year)

# Create an empty list to store the plots
plot_list <- list()


# Loop through unique years and create a plot for each
for (year in unique(result$sale_year)) {
  year_data <- subset(result, sale_year == year)
  
  # Create a horizontal bar plot using ggplot2
  p <- ggplot(year_data, aes(x = total_sales, y = reorder(rep_name, total_sales), fill = rep_name, text = paste("Rep: ", rep_name, "<br>Total Sales : $", total_sales))) +
    geom_bar(stat = "identity") +
    labs(title = paste("Top 5 Sales Reps in", year), x = "Total Sales ", y = "") +
    theme(legend.position = "none") +
    scale_x_continuous(name = "Top 5 Sales Reps", labels = scales::dollar)
  
  # Convert ggplot to plotly
  plot <- ggplotly(p, tooltip = "text")
  
  # Check if the plot for the current year already exists in the list
  if (!(year %in% names(plot_list))) {
    # If not, add the plot to the list
    plot_list[[year]] <- plot
  }
}

# Print the unique plots
print(plot_list)

```

## Analytical Query II
### Total sold per product per quarter

The study provides important insights into the performance of the product throughout the year by revealing the total sales per product for each quarter. For instance, sales of "Zalofen" spiked in 2020's first quarter, but sales of "Bhikatervizem" increased steadily throughout the course of the entire quarter. Businesses can make well-informed decisions about product development, marketing tactics, and inventory management by keeping an eye on these trends. Recognizing the variations in demand every quarter helps businesses make the most use of their resources and take advantage of growth prospects in particular product categories.


```{sql connection=con, label= "Total sold per product per quarter sql", eval = T, warning = F, echo=FALSE, message= FALSE}
SELECT
    product_name,
    sale_year,
    quarter,
    SUM(total_amount_sold) AS total_sold
FROM
    product_facts
GROUP BY
    product_name, sale_year, quarter;
```


```{r label= "Total sold per product per quarter", eval = T, warning = F, echo=FALSE, message= FALSE}

# Check if ggplot2 package is installed, if not, install it
if (!require("ggplot2", quietly = TRUE)) {
  install.packages("ggplot2")
}

# Check if plotly package is installed, if not, install it
if (!require("plotly", quietly = TRUE)) {
  install.packages("plotly")
}

library(ggplot2)
library(plotly)

# Execute the SQL query to retrieve the total sold for each product_name, sale_year, and quarter
query <- "
SELECT
    product_name,
    sale_year,
    quarter,
    SUM(total_amount_sold) AS total_sold
FROM
    product_facts
GROUP BY
    product_name, sale_year, quarter;
"

# Execute the query and store the result in query_result
query_result <- dbGetQuery(con, query)

# Get unique years
unique_years <- unique(query_result$sale_year)

# Iterate over unique years
for (year in unique_years) {
  # Subset data for the current year
  year_data <- subset(query_result, sale_year == year)
  
  # Create ggplot object
  gg <- ggplot(year_data, aes(x = quarter, y = total_sold, color = product_name, group = product_name)) +
    geom_line() +
    labs(title = paste("Total Sold by Quarter for Each Product -", year),
         x = "Quarter",
         y = "Total Sold") +
    theme_minimal() +
    scale_y_continuous(name = "", labels = scales::dollar)
  
  # Convert ggplot to plotly for interactive plot values
  ggplotly(gg, tooltip = "all") %>%
    print()
}


```

## Analytical Query III
### Number of units sold per product per region

Essential information about regional sales trends can be obtained by examining the quantity of units sold of each product in various locations. Businesses can effectively allocate resources and customize marketing tactics to target certain client segments by watching these variables and identifying places with high demand for particular items. For example, "Colophrazen" might sell well in the EMEA region, but "Colophrazen" doesn't do better in the South America area. Companies can improve their distribution networks, inventory control, and product offers to maximize profitability and customer happiness by taking into account these geographical variations.


```{sql connection=con, label= "Number of units sold per product per region sql", eval = T, warning = F, echo=FALSE, message= FALSE}

SELECT
    product_name,
    territory,
    SUM(total_units_sold_per_region) AS total_units_sold
FROM
    product_facts
GROUP BY
    product_name, territory;
    
```



```{r label= "Number of units sold per product per region", eval = T, warning = F, echo=FALSE, message= FALSE}

# Check if ggplot2 package is installed, if not, install it
if (!require("ggplot2", quietly = TRUE)) {
  install.packages("ggplot2")
}

# Check if plotly package is installed, if not, install it
if (!require("plotly", quietly = TRUE)) {
  install.packages("plotly")
}

library(ggplot2)
library(plotly)

# Execute the SQL query to retrieve the number of units sold per product per region
query <- "
SELECT
    product_name,
    territory,
    SUM(total_units_sold_per_region) AS total_units_sold
FROM
    product_facts
GROUP BY
    product_name, territory;
"

# Execute the query and store the result in query_result
query_result <- dbGetQuery(con, query)

# Assuming query_result contains the result of the SQL query
gg <- ggplot(query_result, aes(x = territory, y = total_units_sold, group = product_name, color = product_name)) +
  geom_line() +
  labs(title = "Number of Units Sold per Product per Region",
       x = "Region",
       y = "Total Units Sold $") +
  theme_minimal() +
  scale_y_continuous(name = " ", labels = scales::dollar)

# Convert ggplot to plotly
plotly_gg <- ggplotly(gg, tooltip = "y")

# Print the plot
plotly_gg


```

## Analytical Query IV
### Average sales per sales rep over the years

Examining the mean sales amount per sales agent across time provides significant understanding of individual performance patterns and the efficacy of the sales force as a whole. Businesses can find high-performing sales representatives, spot patterns in sales performance over time, and identify areas for training or improvement by monitoring this indicator. To disseminate best practices throughout the organization, mentorship programs can be informed, for example, by identifying sales representatives who consistently have high average sales. Furthermore, by tracking variations in average sales over time, one may determine how training programs or changes in the market environment affect sales performance, which can help with resource allocation and strategic decision-making. 2020 saw Walison da Silva lead the group, closely followed by Aneeta Kappoorthy and Lynette McRowe. In 2021 and 2022, the pattern persisted, with reliable performers topping the charts.

```{r label= "Average sales per sales rep over the years 1", eval = T, warning = F, echo=FALSE, message= FALSE}
# Load the knitr package
library(knitr)

# SQL query
query <- "
    SELECT 
        rep_name,
        sale_year,
        SUM(total_amount_sold) AS total_sales,
        COUNT(*) AS number_of_sales,
        SUM(total_amount_sold) / COUNT(*) AS average_sales
    FROM 
        rep_facts
    GROUP BY 
        rep_name, sale_year
    ORDER BY 
        rep_name, sale_year;
"

# Execute the query and store the result in 'result'
result <- DBI::dbGetQuery(con, query)

# Display the result using the kable function
kable(result, caption = "Average Sales per Sales Rep Over the Years", align = "c")

```


```{r label= "Average sales per sales rep over the years", eval = T, warning = F, echo=FALSE, message= FALSE}
# Check if ggplot2 package is installed, if not, install it
if (!require("ggplot2", quietly = TRUE)) {
  install.packages("ggplot2")
}

# Check if plotly package is installed, if not, install it
if (!require("plotly", quietly = TRUE)) {
  install.packages("plotly")
}

library(ggplot2)
library(DBI)
library(plotly)

result <- DBI::dbGetQuery(con, "
    SELECT 
        rep_name,
        sale_year,
        SUM(total_amount_sold) AS total_sales,
        COUNT(*) AS number_of_sales,
        SUM(total_amount_sold) / COUNT(*) AS average_sales
    FROM 
        rep_facts
    GROUP BY 
        rep_name, sale_year
    ORDER BY 
        rep_name, sale_year;
")

# Convert sale_year to factor to maintain order on x-axis
result$sale_year <- factor(result$sale_year)
str(df)


# Convert the ggplot object to a plotly object
plotly_plot <- ggplotly(ggplot(data = result, aes(x = sale_year, y = average_sales, color = rep_name, group = rep_name)) +
                          geom_line() +
                          geom_point() +
                          labs(title = "Average Sales per Sales Rep Over the Years",
                               x = "Year",
                               y = "Average Sales ($)",
                               color = "Sales Rep") +
                          theme_minimal())

# Show the interactive plot
plotly_plot

dbDisconnect(con)

```


